---
title: "Regression Models Course Project"
author: "Andrew Mendonca"
date: "July 28, 2017"
output: pdf_document
---

## Executive Summary
Motor Trend is a magazine about the automobile industry. They are interested in 
exploring the relationship between a set of variables and miles per gallon (MPG) (outcome).
They want to know if an automatic or manual is transmission better for MPG and the MPG difference 
between automatic and manual transmissions. In this report, we will evaluate the mtcars data 
set and develop an analysis to answer these questions based on regression models and exploratory 
data analyses.

## Exploratory Data Analysis
Load the following library and data set.

```{r, echo = TRUE}
library(ggplot2)
data(mtcars)
```

View a sample of the data set.
```{r, echo = TRUE}
head(mtcars)
```
Convert variables into factors.
```{r, echo = TRUE}
mtcars$cyl <- factor(mtcars$cyl)
mtcars$vs <- factor(mtcars$vs)
mtcars$am <- factor(mtcars$am, labels = cbind("Automatic", "Manual"))
mtcars$gear <- factor(mtcars$gear)
mtcars$carb <- factor(mtcars$carb)
```
We create exploratory plots to help us understand the data better. Appendix - Plot 1,
illustrates the automatic transmissions having a low MPG than the manual transmissions.

## Regression Analysis
View the difference between automatic and manual transmissions.

```{r, echo = TRUE}
aggregate(mpg ~ am, data = mtcars, mean)
```
We create a hypothesis that automatic cars have a 7.25 MPG lower than manual cars. We then
use a t-test.

```{r, echo = TRUE}
t.test(mtcars[mtcars$am == "Automatic",]$mpg, mtcars[mtcars$am == "Manual",]$mpg)
```
It is shown that the p-value is 0.001374, which is a significant difference. Let's quantify this.

```{r, echo = TRUE}
summary(lm(mpg ~ am, data = mtcars))
```
The average MPG for an automatic transmission is 17.1 MPG, while the manual transmission is
7.2 MPG higher. The R-squared value is 0.36, which means this model only explains 36% of the
variance, so we need a multivariable model.

Appendix - Plot 2, shows how all the other variables correlate with `mpg`. We see that
`cyl`, `disp`, `hp`, and `wt` have strongest correlations with mpg. We include these variables
in the new model and compare them with the initial model.

```{r, echo = TRUE}
anova(lm(mpg~am, data = mtcars), lm(mpg~am + cyl + disp + hp + wt, data = mtcars))
```
The p-value is 8.637e-08, so this multivariable model is significantly better than the linear
model. Appendix - Plot 3, checks the residuals for non-normality and shows that they are all
normally distributed.

```{r, echo = TRUE}
summary(lm(mpg~am + cyl + disp + hp + wt, data = mtcars))
```
This model only explains 86.64% of the variance, so this means that `cyl`, `disp`, `hp`, and
`wt` had an effect on the correlation between `mpg` and `am`. Hence, the difference between
automatic transmissions and manual transmissions is 1.81 MPG.

## Appendix
Plot 1 - Boxplot of MPG based on transmission type

```{r, echo = TRUE}
boxplot(mpg ~ am, data = mtcars, col = (c("green", "orange")), xlab = "Transmission Types",
        ylab = "Miles Per Gallon")
```

Plot 2 - Pairs plot for the rest of the data set

```{r, echo = TRUE}
pairs(mpg ~ ., data = mtcars)
```

Plot 3 - Checking each residual

```{r, echo = TRUE}
par(mfrow = cbind(2,2))
plot(lm(mpg~am + cyl + disp + hp + wt, data = mtcars))
```